### 2014-03-28 paper/discussion.tex
notes from tovi
add figure about voxelization error, reorganize a whole bunch of stuff, basically this is what you will need to work from this weekend!

Once objects have been designed and fabricated, the major flexibility for prototyping with pipes stems from the media inserted in them and the systems attached which sense changes in, or actuate, those media.  The five types of media--gas, liquid, solid, particulate, and threadable--give rise to unique possibilities in output (visual, aural, tactile/haptic, olfactory/gustatory) as well as input (touch/pressure and others).

\valkyrie{this section is probably my project for the weekend, or something.  I can't figure out how to improve it right now.  Tovi, we should talk about this in our meeting this afternoon.}

@@

By crossing pipe types, topologies, and inserted media, we create a space of possible inputs and outputs that can be pipe-mediated.  We offer a table (see Figure \ref{fig:designspace}) containing these possibilities, referencing previous work on input/output techniques where appropriate, noting that many of these techniques have not been attempted using digital fabrication \bjoern{It may be a bit tricky to define what "using digital fab" means here. For example, Harrison 's pneumatically actuated buttons were done on a laser cutter, but also required manual gluing of other layers. Does that count or not?}.  We suggest new points in the design space that have been unexplored or not yet explored with digital fabrication.

\begin{figure*}[t]
\centering
    \includegraphics[width=\textwidth]{figures/designspace.pdf}
\caption{The design space of pipe-based interactions.  Existing systems are written in regular font.  Those created with fabrication are in {\color{blue}blue}. In \emph{{\color{red}red italic}} are unexplored interactions creatable with custom-fabricated pipes. \tovi{Im not sure how well this table will work. Division between input and output should be more prominent.} \valkyrie{one possibility is making the table larger and adding some short descriptions.  we could throw out a lot of words from the paper that way}}
\label{fig:designspace}
\end{figure*}

@@

We describe potential pipe-based outputs organized by the five senses: visual, aural, haptic, and olfactory/gustatory.  In addition to outputs we mention explicitly, any outputs possible using traditional electronic components are possible as described above.

@@

\emph{Visual} outputs can be mediated by gases, liquids, solid, particulates, or threadables.  Pneumatic actuation (as in \cite{Yao-pneui} or \cite{Harrison-buttons}) provides output, but gases carrying smoke (as in \cite{Sodhi-aireal} demo) also offers visual feedback.  Splitting and mixing of colored liquids can be used as a display tactic; additionally mechanical motion can be fluid-mediated.  Liquids can also be mixed in specific concentrations by servo-controlled gate structures as in Figure \ref{fig:direct}.  Printed optics \cite{Willis-printedoptics} explored visual output at the ends of pipes, and   EL wire can be threaded through custom paths to create neon signs, as in Figure \ref{fig:neon}.

@@

\emph{Haptic} outputs are possible through compressible and incompressible fluids.  Fluids can actuate particulates and create programmable pliability \cite{Follmer-jamming}.  Additionally, use of semi-closed pipes with rubberlike material as the caps allows for gases to actuate surface features (see Figure \ref{fig:breathe}).  Free-air haptic feedback \cite{Sodhi-aireal} and volumetric interactions \cite{Iwata-volflex} can be pipe-mediated.

@@

\emph{Touch} sensing can be enabled through pipes filled with conductive media.  Traditional wires (threadable) may be used, but conductive paint (liquid) may be more flexible.  Using conductive paint and Swept Frequency Capacitive Sensing (SFCS \cite{Sato-touche}), an interior star pipe topology can enable single-wire touch- and grasp-sensing at any set of points on a printed object's surface (see Figure \cite{fig:toys}).  Simple capacitance measurements are possible using the same media, and custom-designed sensors like those created in Savage, et al.,'s Midas system are also possible \cite{Savage-midas}.

@@

\emph{Grasp} sensing can be enabled using Wimmer's FlyEye techniques \cite{Wimmer-flyeye} or those described in Jamming User Interfaces \cite{Follmer-jamming}.  The optical links necessary for sensing can be created either via clear solid cores as in \cite{Willis-printedoptics} or via fiber optic cables threaded through hollow pipes (see Figure \ref{fig:pens}).

@@

\bjoern{The biggest question wrt fabrication for me is - how do I print these cavities? So general support material avoidance or removal strategies should be more prominent.

(I think the lye bath would probably work better than the high pressure water for our objet prints)}
The physical fabrication process used by different machines gives rise to differing design guidelines.  The two machines we used to create our example objects were the Objet Connex 260 and the Makerbot Replicator 2.

The Objet is an inkjet-based 3D printer which jets drops of polymer in very fine layers, curing each layer with an ultraviolet lamp.  It offers a wide range of material types, including clear/opaque, flexible/hard, and black/white.  Additionally, it permits the mixing of two different materials in a single print: a mixed material could include 70\% flexible and 30\% hard, for example, to be firm but somewhat flexible.  The flexible materials can be inflated when they are thin: they elongate 200\% before breaking.  We have found $.05$inches to be a good thickness for inflatable materials.  Although this machine has many extraordinary capabilities, it unfortunately cannot print overhangs of greater than $14^{\circ}$ without support material.  The support structure used by this machine is a loose matrix of material that can be removed by hand from the surface or blasted out with a water jet from the interior.  The support is also soluble in lye.

\subsection{Identity}
Pipes included in objects can have additional effects beyond their designed intentions.

Pipes included in printed objects allow for their identification.  While the pipes need not be visible (especially in the case of fully-enclosed pipes), their presence, location, network structure, and length change the physical properties of a printed object.  This includes weight, acoustic resonance, and (if conductive material is present) capacitive signature.

@@

>>>>

\tovi{we should make this into a "1D table", just base it on the I/O we can enable.  scrap the table illustration itself; it's not needed.  also, focus more on future work in these areas, and work on relating back to what we've discussed already (and the examples we've built)}

@@

By crossing pipe types, topologies, and inserted media, we create a space of possible inputs and outputs that can be pipe-mediated.  We arrange our I/O discussion around the modes of input or output (the five senses, touch/tapping/pressure/etc.).  We also suggest new points in the design space of pipe-mediated interaction that remain unexplored.

@@

We describe potential pipe-based outputs organized by the five senses: visual, aural, haptic, and olfactory/gustatory.  In addition to outputs we mention explicitly, any outputs possible using traditional electronic components are possible using conductive material in pipes.

@@

\emph{Visual} outputs can be mediated by gases, liquids, solid, particulates, or threadables.  Pneumatic actuation (as in PneUIs \cite{Yao-pneui} or Harrison's latex buttons \cite{Harrison-buttons}) provides output, but gases carrying smoke (as in the Aireal \cite{Sodhi-aireal} demo) also offers visual feedback.  Splitting and mixing of colored liquids can be used as a display tactic; additionally mechanical motion can be fluid-mediated.  Liquids can also be mixed in specific concentrations by servo-controlled gate structures as in Figure \ref{fig:direct}.  Printed optics \cite{Willis-printedoptics} explored visual output at the ends of pipes, and  EL wire can be threaded through custom paths to create neon signs, as in our neon robot (Figure \ref{fig:neon}).

@@

\emph{Haptic} outputs are possible through compressible and incompressible fluids.  Fluids can actuate particulates and create programmable pliability \cite{Follmer-jamming}.  Additionally, use of semi-closed pipes with rubberlike material as the caps allows for gases to actuate surface features, like the back of our breathing rabbit (see Figure \ref{fig:breathe}).  Free-air haptic feedback as in Aireal \cite{Sodhi-aireal} and volumetric interactions as in Volflex \cite{Iwata-volflex} can also be pipe-mediated.

@@

\emph{Touch} sensing can be enabled through pipes filled with conductive media.  Traditional wires (threadable) may be used, but conductive paint (liquid) may be more flexible.  Using conductive paint and Swept Frequency Capacitive Sensing (SFCS \cite{Sato-touche}), an interior star pipe topology can enable single-wire touch- and grasp-sensing at any set of points on a printed object's surface, like our toys (see Figure \cite{fig:toys}).  Simple capacitance measurements are possible using the same media, and custom-designed sensors like those created in Savage, et al.,'s Midas system are also possible \cite{Savage-midas}.

@@

\emph{Grasp} sensing can be enabled using Wimmer's FlyEye techniques \cite{Wimmer-flyeye} or those described in Jamming User Interfaces \cite{Follmer-jamming}.  The optical links necessary for sensing can be created either via clear solid cores as in Printed Optics \cite{Willis-printedoptics} or via fiber optic cables threaded through hollow pipes (see Figure \ref{fig:pens}).

@@

\subsubsection{Identification}
In addition to active inputs from users, pipes can be used for passive \emph{identification}.  While the pipes need not be visible (especially in the case of fully-enclosed pipes), their presence, location, network structure, and length change the physical properties of a printed object.  This includes weight, acoustic resonance, and (if conductive material is present) capacitive signature.

@@

\bjoern{The biggest question wrt fabrication for me is - how do I print these cavities? So general support material avoidance or removal strategies should be more prominent.)}

The physical fabrication process used by different machines gives rise to differing strategies for avoiding support material or easing its removal.  The two machines we used to create our example objects were the Objet Connex 260 and the Makerbot Replicator 2.

The Objet is an inkjet-based 3D printer which jets drops of polymer in very fine layers, curing each layer with an ultraviolet lamp.  It offers a wide range of material types, including clear/opaque, flexible/hard, and black/white.  Additionally, it permits the mixing of two different materials in a single print: a mixed material could include 70\% flexible and 30\% hard, for example, to be firm but somewhat flexible.  The flexible materials can be inflated when they are thin: they elongate 200\% before breaking.  We have found $.05$ inches to be a good thickness for inflatable materials.  Although this machine has many extraordinary capabilities, it unfortunately cannot print overhangs of greater than $14^{\circ}$ without support material.  Due to this, support avoidance is difficult; our main tactic for removing support material has been to make the \emph{openings} to pipes as large as possible.  This allows us to  \valkyrie{you are here!}  The support structure used by this machine is a loose matrix of material that can be removed by hand from the surface, blasted out with a water jet from large cavities, or dissolved in lye (the enclosing model material does not dissolve).
